---
title: "categories_project"
author: "Haichuan Du"
date: "4/17/2020"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(dplyr) # Lirary for spliting train & test dataset
library(corrplot) # Plotting nice correlation matrix
library(caret)
library(pROC) # For checking ROC Curve of the model
library(ROCR)
library(MASS)
library(bestglm)
library(brglm)
library(psych)
library(ResourceSelection)
library(lmtest)
library(gmodels)
library(mice)
library(nnet)
library(VIM)
library(glmnet)
```

```{r}
setwd("/Users/haidu/Desktop/statistics/Categorical\ Data\ Analysis\ /project")
loan <- read.csv("loan.csv", header=T, na.strings=c("","NA"))
loan_data <- loan
attach(loan_data)
loan_data <- dplyr::select(loan_data, -Loan_ID) #Removing Loan_ID as it has no logical corelation
head(loan_data)
```

```{r}
sum(is.na(loan_data))  ## Checking for total missing values
colSums(is.na(loan_data)) ##  checking for any missing values in the feature
aggr(loan,prop=FALSE,numbers=TRUE)
```

## Explore data visually
```{r}
ggplot(data=loan_data) +
  geom_point(aes(x=LoanAmount, y=Loan_Status, color=Property_Area)) 

ggplot(data=loan_data) +
  geom_bar(aes(x=Loan_Status,color=Loan_Status)) 

ggplot(data=loan_data) +
  geom_bar(aes(x=Loan_Amount_Term, fill=Loan_Status ))

ggplot(data=loan_data) +
  geom_histogram(aes(x=LoanAmount), bins = 50)

ggplot(data=loan_data) +
  geom_histogram(aes(x=ApplicantIncome))

ggplot(data=loan_data) +
  geom_histogram(aes(x=CoapplicantIncome ))

ggplot(data=loan_data) +
  geom_bar(aes(x=Credit_History, fill=Loan_Status))

ggplot(data=loan_data) +
  geom_bar(aes(x=Dependents,fill=Loan_Status ))

ggplot(data=loan_data) +
  geom_bar(aes(x=Education, fill=Loan_Status))

ggplot(data=loan_data) +
  geom_bar(aes(x=Married, fill=Loan_Status))

#Making Contingency Table to check percentage of Credit_History in relation with  loan status
CrossTable(loan_data$Loan_Status, loan_data$Credit_History,prop.r = TRUE, prop.c = FALSE, prop.t = FALSE,
           prop.chisq = FALSE)

CrossTable(loan_data$Loan_Status, loan_data$Married,prop.r = TRUE, prop.c = FALSE, prop.t = FALSE,
           prop.chisq = FALSE )

CrossTable(loan_data$Loan_Status, loan_data$Education,prop.r = TRUE, prop.c = FALSE, prop.t = FALSE,
           prop.chisq = FALSE )

CrossTable(loan_data$Loan_Status, loan_data$Self_Employed,prop.r = TRUE, prop.c = FALSE, prop.t = FALSE,
           prop.chisq = FALSE )

CrossTable(loan_data$Loan_Status, loan_data$Property_Area,prop.r = TRUE, prop.c = FALSE, prop.t = FALSE,
           prop.chisq = FALSE )
```

## Handling Null, Missing and Categorical Variables
inputs the mean into the missing values, similary for categorical variable, we can use the category that apperars the most frequently
```{r}
loan_data <- loan_data %>% 
  mutate(LoanAmount=ifelse(is.na(LoanAmount), mean(LoanAmount, na.rm = T), LoanAmount),
         Loan_Amount_Term=ifelse(is.na(Loan_Amount_Term), median(Loan_Amount_Term, na.rm = T), Loan_Amount_Term),
         Credit_History=ifelse(is.na(Credit_History), 1, Credit_History))
```

## Handling  Outlier
(Replace outlier with lower and upper cutoff value)
using the rule of thumb where upper limit is computed as 1.5 * IRQ, where IRQ = 3rd Quartile – 1st Quartile. 
```{r}
#scatter plot  to detect outliers for ApplicantIncome
plot(ApplicantIncome, ylab = "ApplicantIncome")

outliers_upperlimit_AppIncome <- quantile(ApplicantIncome, 0.75) + 1.5 * IQR(ApplicantIncome) # upper_limit = 10171.25
index.outliers.ApplicantIncome <- which(ApplicantIncome > outliers_upperlimit_AppIncome | ApplicantIncome < 0 ) # 50 outliers
loan_data <- loan_data[-index.outliers.ApplicantIncome,] #Removing observations
plot(loan_data$ApplicantIncome, ylab = "ApplicantIncome") 

# detect outliers for CoapplicantIncome
plot(loan_data$CoapplicantIncome, ylab = "CoapplicantIncome")
outliers_upperlimit_CoIncome <- quantile(loan_data$CoapplicantIncome, 0.75) + 1.5 * IQR(loan_data$CoapplicantIncome) 
index.outliers.CoIncome <- which(loan_data$CoapplicantIncome > outliers_upperlimit_CoIncome | loan_data$CoapplicantIncome < 0 ) 
loan_data <- loan_data[-index.outliers.CoIncome,] #Removing observations
plot(loan_data$CoapplicantIncome, ylab = "CoapplicantIncome")  

# Treatment of outlier for LoanAmount
plot(loan_data$LoanAmount, ylab = "LoanAmount")
outliers_upperlimit_LoanAmount <- quantile(loan_data$LoanAmount, 0.75) + 1.5 * IQR(loan_data$LoanAmount) 
index.outliers.LoanAmount <- which(loan_data$LoanAmount > outliers_upperlimit_LoanAmount | loan_data$LoanAmount < 0 ) 
loan_data <- loan_data[-index.outliers.LoanAmount,] #Removing observations
plot(loan_data$LoanAmount, ylab = "LoanAmount")  
```

## Transform the categorical data
create dummy variables for categorical attributes 
```{r}
#I converted the Dependents variable to a continuous variable in order
loan_data$Dependents=as.numeric(substr(loan_data$Dependents,1,1)) 

loan_data <- loan_data %>%
  mutate(Gender=ifelse(Gender=="Male",1,0),
         Married=ifelse(Married=="Yes",1,0),
         Education=ifelse(Education=="Graduate",1,0),
         Self_Employed=ifelse(Self_Employed=="Yes",1,0),
         Loan_Status=ifelse(Loan_Status=="Y",1,0))

#deal with missing value again, for catergorical data, use the category that appears most frequently
loan_data <- loan_data %>%
  mutate(Gender=ifelse(is.na(Gender),1,Gender),
         Married=ifelse(is.na(Married),1,Married),
         Dependents=ifelse(is.na(Dependents),0,Dependents),
         Self_Employed=ifelse(is.na(Self_Employed),0,Self_Employed))

# More than 2 unique values treatment
loan_data$Urban <- ifelse(loan_data$Property_Area=="Urban",1,0)
loan_data$Rural <- ifelse(loan_data$Property_Area=="Rural",1,0)
loan_data$Semiurban <-ifelse(loan_data$Property_Area=="Semiurban",1,0)

```

#check for correlation between the variables
exploratory data analysis is to check correlations among all variables 
```{r}
cor(loan_data[, sapply(loan_data, class) != "factor" ],) #Checking multicollinearity 

# graph check multicollinearity
correlation <- loan_data[sapply(loan_data, is.numeric)]
descrCorr <- cor(correlation)
corrplot(descrCorr)

pairs.panels(loan_data)
```

## Building Logistic Regression Model1 by using Multiple Imputation dataset
```{r}
# let's take a look initial model
head(loan_data)
loan_data_R <- dplyr::select(loan_data, -Property_Area)
reg1 <- glm(Loan_Status~., family = binomial, data = loan_data_R)
summary(reg1)
cor(loan_data_R$Loan_Status, fitted(reg1))  #0.60
hoslem.test(loan_data_R$Loan_Status, fitted(reg1))  #p-value = 0.4663

####################################################################################

#mice treating missing valued by using the loan data = origional data

loan <- dplyr::select(loan,-Loan_ID)

aggr(loan,prop=FALSE,numbers=TRUE)
aggr(loan,prop=TRUE,numbers=TRUE)

#Multiple Imputation
imps = mice(loan)

library(lattice)
densityplot(imps)

#What we would like to see is that the shape of the magenta points (imputed) matches the shape of the blue ones (observed). The matching shape tells us that the imputed values are indeed “plausible values”.The density of the imputed data for each imputed dataset is showed in magenta while the density of the observed data is showed in blue. Again, under our previous assumptions we expect the distributions to be similar.

#mice reg model
fit_mcie = with(imps, glm(loan$Loan_Status~., family = binomial, data = loan))
Final_micefit <- pool(fit_mcie)
summary(Final_micefit)

# porpuslful selection
#step1
summary(pool(with(imps, glm(loan$Loan_Status~Gender, family = binomial, data = loan))))
summary(pool(with(imps, glm(loan$Loan_Status~Married, family = binomial, data = loan))))
summary(pool(with(imps, glm(loan$Loan_Status~Dependents, family = binomial, data = loan))))
summary(pool(with(imps, glm(loan$Loan_Status~Education, family = binomial, data = loan))))
summary(pool(with(imps, glm(loan$Loan_Status~Self_Employed, family = binomial, data = loan))))
summary(pool(with(imps, glm(loan$Loan_Status~ApplicantIncome, family = binomial, data = loan))))
summary(pool(with(imps, glm(loan$Loan_Status~CoapplicantIncome, family = binomial, data = loan))))
summary(pool(with(imps, glm(loan$Loan_Status~LoanAmount, family = binomial, data = loan))))
summary(pool(with(imps, glm(loan$Loan_Status~Loan_Amount_Term, family = binomial, data = loan))))
summary(pool(with(imps, glm(loan$Loan_Status~Credit_History, family = binomial, data = loan))))
summary(pool(with(imps, glm(loan$Loan_Status~Property_Area, family = binomial, data = loan))))
#we find that Married， Education， CoapplicantIncome, and Credit_History have p-values less than 0.2. 

summary(pool(with(imps, glm(loan$Loan_Status~Married+Education+CoapplicantIncome
                            +Credit_History, family = binomial, data = loan))))

#step2
summary(pool(with(imps, glm(loan$Loan_Status~Married+Education+Credit_History, family = binomial, data = loan))))

summary(pool(with(imps, glm(loan$Loan_Status~Married+Credit_History, family = binomial, data = loan))))


#step3 
summary(pool(with(imps, glm(loan$Loan_Status~Married+Credit_History       
                            +Dependents, family = binomial, data = loan))))

summary(pool(with(imps, glm(loan$Loan_Status~Credit_History+Married     
                            +Education, family = binomial, data = loan))))

summary(pool(with(imps, glm(loan$Loan_Status~Credit_History+Married     
                            +Self_Employed, family = binomial, data = loan))))

summary(pool(with(imps, glm(loan$Loan_Status~Credit_History+Married     
                            +ApplicantIncome, family = binomial, data = loan))))

summary(pool(with(imps, glm(loan$Loan_Status~Credit_History+Married     
                            +CoapplicantIncome, family = binomial, data = loan))))

summary(pool(with(imps, glm(loan$Loan_Status~Credit_History+Married     
                            +LoanAmount, family = binomial, data = loan))))       # only Loan_amount change Marries by at 10%

summary(pool(with(imps, glm(loan$Loan_Status~Credit_History+Married     
                            +Loan_Amount_Term, family = binomial, data = loan))))

summary(pool(with(imps, glm(loan$Loan_Status~Credit_History+Married     
                            +Credit_History, family = binomial, data = loan))))

summary(pool(with(imps, glm(loan$Loan_Status~Credit_History+Married     
                            +Property_Area, family = binomial, data = loan))))


#step4:Attempt adding plausible interactions among variables in the model, usually using somewhat stricter standards such a p-value<0.05 (can consider non-linear predictor terms, like quadratic effects, in this step as well).

summary(pool(with(imps, glm(loan$Loan_Status~Credit_History+Married     
                            +LoanAmount + Credit_History:Married, family = binomial, data = loan))))   

summary(pool(with(imps, glm(loan$Loan_Status~Credit_History+Married     
                            +LoanAmount + Married:LoanAmount, family = binomial, data = loan))))    

summary(pool(with(imps, glm(loan$Loan_Status~Credit_History+Married     
                            +LoanAmount + Credit_History:LoanAmount,
                            family = binomial, data = loan))))     # Credit_History:LoanAmount p = 0.034

summary(pool(with(imps, glm(loan$Loan_Status~Credit_History+Married     
                            +LoanAmount+Credit_History:LoanAmount+I(LoanAmount^2), family = binomial, data = loan))))     # Credit_History:LoanAmount p = 0.046

# Conclusion：model included: Credit_History,  Married, LoanAmount,  Credit_History*LoanAmount


########################################################################
#complete the missing value by using final model predictor see what's going on
one_of_dataset <- complete(imps,1)
model1 <-glm(Loan_Status~Credit_History+Married+LoanAmount+
                        Credit_History:LoanAmount, family = binomial, data = one_of_dataset)

rocplotone11 <- roc(one_of_dataset$Loan_Status ~ fitted(model1), data = one_of_dataset)
plot.roc(rocplotone11, print.auc = TRUE)
auc(rocplotone11)      #0.78 Auc depdent on different dataset


#using different complete dataset and stepAIC() see what's going on, 
stepAIC(glm(Loan_Status~., family = binomial, data = one_of_dataset))

#run the smallest AIC
aci_model<-glm(formula = Loan_Status ~ Married + Credit_History + Property_Area, 
    family = binomial, data = one_of_dataset)

#ROC
rocplotone <- roc(one_of_dataset$Loan_Status ~ fitted(aci_model), data = one_of_dataset)
plot.roc(rocplotone, print.auc = TRUE)
auc(rocplotone)

```

## Building Logistic Regression Model2 withdataset that replaced by using the mean of each feature separately
```{r}
#Purposeful Model Selection  
##step1 Fit “simple” logistic regression models for each of the predictors separately.
#Eliminate any predictor values with large p-values (say >0.2).

summary(glm(Loan_Status ~ Gender, family=binomial, data = loan_data_R))
summary(glm(Loan_Status ~ Married, family=binomial, data = loan_data_R))
summary(glm(Loan_Status ~ Dependents, family=binomial, data = loan_data_R))
summary(glm(Loan_Status ~ Education, family=binomial, data = loan_data_R))
summary(glm(Loan_Status ~ Self_Employed, family=binomial, data = loan_data_R))
summary(glm(Loan_Status ~ ApplicantIncome, family=binomial, data = loan_data_R))
summary(glm(Loan_Status ~ CoapplicantIncome, family=binomial, data = loan_data_R))
summary(glm(Loan_Status ~ LoanAmount, family=binomial, data = loan_data_R))
summary(glm(Loan_Status ~ Loan_Amount_Term, family=binomial, data = loan_data_R))
summary(glm(Loan_Status ~ Credit_History, family=binomial, data = loan_data_R))
summary(glm(Loan_Status ~ Urban, family=binomial, data = loan_data_R))
summary(glm(Loan_Status ~ Rural, family=binomial, data = loan_data_R))

summary(glm(formula=Loan_Status ~ Married + Education + CoapplicantIncome + Credit_History + Rural, family=binomial, data = loan_data_R))

#step2 Conduct forward stepwise selection with remaining predictors, usually using a more stringent cut-off, such as p-value<0.1 or perhaps AIC/BIC. 

summary(glm(formula=Loan_Status ~ Married + Education + Credit_History + Rural, family=binomial, data=loan_data_R))


#step3 Consider adding in any variables that were not included in the model after Step 1 or Step 2.  A predictor can be added in even if p-value>0.1 if the AIC/BIC is lower or if it changes the estimated β coefficients by at least, say, 10%.

summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + Urban, family=binomial, data = loan_data_R))            # chnage rural by at 50%

summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + Semiurban, family=binomial, data = loan_data_R))        


summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + Loan_Amount_Term, family=binomial, data = loan_data_R))        


summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + LoanAmount, family=binomial, data = loan_data_R))  

summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + CoapplicantIncome, family=binomial, data = loan_data_R))        #change married by aroudn 18%

summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + ApplicantIncome, family=binomial, data = loan_data_R))      

summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + Self_Employed, family=binomial, data = loan_data_R))     

summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + Dependents, family=binomial, data = loan_data_R))        


# we adding back semiurban first


summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + Urban, family=binomial, data = loan_data_R))        # chnage rural by more than 50%


summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + Urban + CoapplicantIncome, family=binomial, data = loan_data_R))       #chnage married about 17% and rural about 30%


summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + Urban + CoapplicantIncome + LoanAmount, family=binomial, data = loan_data_R))       #chnage over 10%


summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + Urban + CoapplicantIncome + LoanAmount + Dependents, family=binomial, data = loan_data_R))      


summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + Urban + CoapplicantIncome + LoanAmount + ApplicantIncome, family=binomial, data = loan_data_R))     

#the following are just roughtly change near 10% but p valur is very large we just ingore

summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + Urban + CoapplicantIncome + LoanAmount + Self_Employed, family=binomial, data = loan_data_R))      


#4Attempt adding plausible interactions among variables in the model,considering interactions as follows does not produce anything extra

summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + Urban + CoapplicantIncome + LoanAmount + LoanAmount:Credit_History, family=binomial, data = loan_data_R))       


summary(glm(formula=Loan_Status ~ Married + Education + Credit_History 
            + Rural + Urban + CoapplicantIncome + LoanAmount + LoanAmount:Education, family=binomial, data = loan_data_R))    

summary(glm(formula=Loan_Status ~ Married + Education + Credit_History
            + Rural+ Urban + LoanAmount +CoapplicantIncome + Married:Credit_History, family=binomial, data = loan_data_R))

summary(glm(formula=Loan_Status ~ Married + Education + Credit_History
            + Rural+ Urban + LoanAmount +CoapplicantIncome + LoanAmount:CoapplicantIncome, family=binomial, data = loan_data_R))


# thus, the finla model included these presictors
# Married, Credit_History, Urban, Rural,Education, CoapplicantIncome, LoanAmount

hand_best_model_dataset <-dplyr::select(loan_data_R, Married, Credit_History, Urban, Rural,Education, CoapplicantIncome, Loan_Status, LoanAmount)

bestModel <- glm(Loan_Status ~ Married + Credit_History + Urban + Rural 
                 + Education + CoapplicantIncome + LoanAmount, family = binomial, data = loan_data_R)
summary(bestModel)


cor(hand_best_model_dataset$Loan_Status, fitted(bestModel)) # R=0.6 is useful for comparing fits of different models for the same data.

#ROC
rocBestModel <- roc(hand_best_model_dataset$Loan_Status ~ fitted(bestModel))
plot.roc(rocBestModel, print.auc = TRUE)
auc(rocBestModel) #0.81

#hoslem.test
hoslem.test(loan_data_R$Loan_Status, fitted(bestModel)) #0.06623


#Confusion Matrix
fit123 = predict(bestModel, newdata = loan_data_R, type = 'response') 
# If results are more than 50% then convert to 1 else 0
fit123 = ifelse(fit123 >=0.5,1,0) #Setting cut-off to be at 0.5
# Evaluate predictions on the training dataset through Confusion Matrix
cnf_matrix = table(predicted = fit123, actual = loan_data_R$Loan_Status)
cnf_matrix
TN = cnf_matrix[1,1] # True Negative - Actual & Predicted is 0/N
TP = cnf_matrix[2,2] # True Positive - Actual & Predicted is 1/Y
FP = cnf_matrix[2,1] # False Positive - Actual is 0/N but Predicted is 1/Y
FN = cnf_matrix[1,2] # False Nefgative - Actual is 1/Y but Predicted is 0/N
TO = TN+TP+FP+FN # Total Observations

accuracy = (TP+TN)/TO # Accuracy or Prevalance of Confusion Matrix = 0.82
accuracy
precision = TP/(TP+FP) # Precision = 0.80
precision
sensitivity = TP/(TP+FN) # True Positive Rate = 0.98
sensitivity
error = (FP+FN)/TO # Error Rate=0.18
error
specificity = TN/(TN+FP)
specificity      #0.45
G=sqrt(specificity*sensitivity) ##G-mean=0.67
G


####################################################################################
#Aautomatic backward selection
summary(reg1)
reg2 <- update(reg1,.~.-Semiurban)
summary(reg2)
stepAIC(reg2, direction = "backward")
```

## Fit the lasso regression model

```{r}
x<-cbind(loan_data_R$Married, loan_data_R$Education, loan_data_R$CoapplicantIncome,
    loan_data_R$LoanAmount, loan_data_R$Credit_History, loan_data_R$Urban, loan_data_R$Rural,
    loan_data_R$Gender, loan_data_R$Dependents, loan_data_R$Self_Employed, loan_data_R$ApplicantIncome, loan_data_R$Loan_Amount_Term +loan_data_R$Semiurban)  # Combine variables by column

y<-loan_data_R$Loan_Status

grid<-10^seq(10,-2, length=100)        # Create a grid of lambda values
lasso.mod=cv.glmnet(x,y,lambda=grid,   # Build a CV lasso regression          
nfold=length(y),                       # nfold=sample size, leave-one-out CV
alpha=1)                               # alpha=0, lasso reg is fit

## Ignore the received warning which recommends leaving 3-or-more out in CV ## 
#Warning message:
#Option grouped=FALSE enforced in cv.glmnet, since < 3 observations per fold 

plot(log10(lasso.mod$lambda), lasso.mod$cvm,      # Plot average CV error versus log(lambda)
xlab="log10(Lambda)", ylab="CV Error")                     
abline(v = log10(lasso.mod$lambda.min), lty = 3)

(lambda=lasso.mod$lambda.min)        # The lambda that minimizes CV error

predict(lasso.mod,s=lambda,          # Obtain lasso reg coefs
type="coefficients")


#Confusion Matrix
fit1234 = predict(lasso.mod, newx = x, type = 'response') 
# If results are more than 50% then convert to 1 else 0
fit1234 = ifelse(fit1234 >=0.5,1,0) #Setting cut-off to be at 0.5
# Evaluate predictions on the training dataset through Confusion Matrix
cnf_matrix2 = table(predicted = fit1234, actual = loan_data_R$Loan_Status)
cnf_matrix2
TN2 = cnf_matrix2[1,1] # True Negative - Actual & Predicted is 0/N
TP2 = cnf_matrix2[2,2] # True Positive - Actual & Predicted is 1/Y
FP2 = cnf_matrix2[2,1] # False Positive - Actual is 0/N but Predicted is 1/Y
FN2 = cnf_matrix2[1,2] # False Nefgative - Actual is 1/Y but Predicted is 0/N
TO2 = TN2+TP2+FP2+FN2 # Total Observations

accuracy2 = (TP2+TN2)/TO2 # Accuracy or Prevalance of Confusion Matrix
accuracy2
precision2 = TP2/(TP2+FP2) 
precision2
sensitivity2 = TP2/(TP2+FN2) 
sensitivity2
error2 = (FP2+FN2)/TO2 
error2
specificity2 = TN2/(TN2+FP2)
specificity2    

#G-mean
G2=sqrt(specificity2*sensitivity2) 
G2

```

